---
title: "Early Diabetes Prediction Tool"
author: "Joel Chavez Gomez"
date: "4/2/2022"
output: 
        ioslides_presentation:
                widescreen: true
---

<link href='http://fonts.googleapis.com/css?family=Baumans' rel='stylesheet' type='text/css'>
<style type="text/css">
.title-slide hgroup > h1{
  font-family: 'Sans-Serif', cursive;
}

<style type="text/css">


.title-slide hgroup > h2{
  font-family: 'Lato', 'Calibri', sans-serif;
}

.title-slide {
  background-color: #e5e5e5; /* Background color for title slide*/
  /* background-image:url(http://goo.gl/EpXln); */
}


.title-slide hgroup > h1, 
.title-slide hgroup > h2 {
  font-family: 'Lato', 'Calibri', sans-serif; /* font for title */
  color: #457b9d /* color for title in title slide */
}

/* Fonts and Spacing */
article p, article li, article li.build, section p, section li{
  font-family: 'Open Sans','Helvetica', 'Crimson Text', 'Garamond',  'Palatino', sans-serif;
  text-align: justify;
  font-size:22px;
  line-height: 1em;
  color: #3e5c76;
}

slide:not(.segue) h2{
  font-family: 'Calibri', Arial, sans-serif;
  font-size: 50px;
  font-style: normal;
  font-weight: bold;
  text-transform: normal;
  letter-spacing: -1px;
  line-height: 1.2em;
/*  color: #193441;*/
/*  color: #02574D;*/
  color: #ee6c4d; /* color for slides header title */
}

/* Reduce Space between Title and Body */
slides > slide > hgroup + article {
  margin-top: 10px;
}

slides > slide {
    background: #ffffff;
}

</style>

```{r setup, include = FALSE, warning=FALSE, message=FALSE}
library(caret)
library(tidyverse)
library(data.table)
source("C:/Users/chave/Desktop/DDP-FinalProject/DevelopingDataProducts/Assets/Script/model_script.R")
```
## Introduction 
- Diabetes is a chronic, metabolic disease characterized by elevated levels of
blood glucose, which leads over time to serious damage to the heart, blood vessels, eyes, kidneys and nerves.  

- The most common is type 2 diabetes, usually in adults, which occurs when the body becomes resistant to insulin or doesn't make enough insulin. 

- About 422 million people worldwide have diabetes, the majority living in low-and middle-income countries, and 1.5 million deaths are directly attributed to diabetes each year.

- Both the number of cases and the prevalence of diabetes have been steadily increasing over the past few decades. 

- Better tools for screening are needed in order to detect early-onset diabetes
and reduce morbidity and mortality.

## Our Predictor Tool
- A *Random Forest* model was created with the [Early Stage Diabetes Risk Prediction Dataset](https://www.kaggle.com/ishandutta/early-stage-diabetes-risk-prediction-dataset?select=diabetes_data_upload.csv) from *kaggle*.

- The [Web App](https://chavezgjoel.shinyapps.io/EarlyDiabetesPrediction/?_ga=2.262725252.901862142.1643921629-813351577.1643921629) has a sidebar panel, in which the user inputs the corresponding patient data. The model takes this input and displays a prediction in the main
panel, along with the probability of being classified as *Positive* or *Negative*.

- Instructions for the use of the App can be found in the [documentation].

- The script to generate the Training and Testing Data sets, the model and it's
corresponding confusion matrix can be found on the [GitHub Repository](https://github.com/JoelChG/DiabetesShinyApp)


The implementation of Machine Learning models tools as Decision Support Systems
for health care providers could help with the screening and early identification
of patients at risk of diabetes or at early stages, before complications occur.

DISCLAIMER: it is important to note that this is not a diagnostic tool, and should not be used to establish or exclude a diagnosis.

## Confusion Matrix
These are the result of the confusion matrix for our model predictions on the testing set.

```{r echo = TRUE}
confusionMat$table
as.matrix(confusionMat$byClass, dimnames = TRUE)[c(1:4, 8), ]
```

## References
<br>
<br>

1. Islam, MM Faniqul, et al. 'Likelihood prediction of diabetes at early stage using data mining techniques.' Computer Vision and Machine Intelligence in Medical Image Analysis. Springer, Singapore, 2020. 113-125.  
2. Diabetes, [World Health Organization](https://www.who.int/health-topics/diabetes#tab=tab_1)

`